---
title: "Analysis of MTT assay results of EDA-treated astrocytes from multiple experiments"
author: "Evgenii O. Tretiakov"
date: "`r Sys.Date()`"
output:
  workflowr::wflow_html:
    toc: yes
    toc_float: yes
    theme: journal
    highlight: textmate
    code_folding: hide
    df_print: paged
editor_options:
  markdown:
    wrap: 72
---

```{r setup, echo=FALSE, include=FALSE}
docname <- "mini-meta-dabest-of-mtt-eda-astrocytes"
now <- Sys.time()
# Time chunks during knitting
knitr::knit_hooks$set(timeit = function(before) {
  if (before) {
    print(paste("Start:", Sys.time()))
    now <<- Sys.time()
  } else {
    print(paste("Stop:", Sys.time()))
    print(Sys.time() - now)
  }
})

knitr::opts_chunk$set(
  autodep        = TRUE,
  cache          = FALSE,
  cache.path     = paste0("cache/", docname, "/"),
  cache.comments = FALSE,
  cache.lazy     = FALSE,
  dev            = c("png", "pdf"),
  echo           = TRUE,
  error          = FALSE,
  fig.align      = "center",
  fig.width      = 14,
  fig.height     = 12,
  message        = FALSE,
  warning        = FALSE,
  timeit         = TRUE
)
# Load packages
suppressPackageStartupMessages({
  library(here)
  library(knitr)
  library(RColorBrewer)
  library(dabestr)
  library(tidyverse)
  library(magrittr)
  library(stringr)
  library(skimr)
})

# Set paths
src_dir <- here("code")
data_dir <- here("data")
output_dir <- here("output")
plots_dir <- here(output_dir, "figures")
tables_dir <- here(output_dir, "tables")

# parallelisation
n_cores <- 4

# set seed
reseed <- 42
set.seed(seed = reseed)

# ggplot2 theme
theme_set(ggmin::theme_powerpoint())
```

```{r load}
df <- read_tsv(here(data_dir, "ASTRO_MTT_EPA.tsv"))
df <- df %>%
  tidyr::gather(key = Group, value = Measurement)

kable(df)
```

```{r load-glut}
df_glut <- read_tsv(here(data_dir, "ASTRO_MTT_Glut_24h.tsv"), col_types = "d")
df_glut <- df_glut %>%
  tidyr::gather(key = Group, value = Measurement)

kable(df_glut)
```

## 5 μM of EPA - Figure S5b

```{r}
unpaired5 <- load(df,
  x = Group, y = Measurement,
  idx = list(
    c("Ctr_1", "EPA_5μM_1"),
    c("Ctr_2", "EPA_5μM_2"),
    c("Ctr_3", "EPA_5μM_3"),
    c("Ctr_4", "EPA_5μM_4"),
    c("Ctr_5", "EPA_5μM_5"),
    c("Ctr_6", "EPA_5μM_6")
  ),
  minimeta = TRUE
)
```

```{r}
print(unpaired5)
```

```{r}
unpaired5.mean_diff <- mean_diff(unpaired5)

print(unpaired5.mean_diff)
```

```{r}
kable(unpaired5.mean_diff$boot_result |> select(-bootstraps))
```

```{r}
dabest_plot(unpaired5.mean_diff)
```

## 10 μM of EPA - Figure S5b1

```{r}
unpaired10 <- load(df,
  x = Group, y = Measurement,
  idx = list(
    c("Ctr_1", "EPA_10μM_1"),
    c("Ctr_2", "EPA_10μM_2"),
    c("Ctr_3", "EPA_10μM_3"),
    c("Ctr_4", "EPA_10μM_4"),
    c("Ctr_5", "EPA_10μM_5"),
    c("Ctr_6", "EPA_10μM_6")
  ),
  minimeta = TRUE
)
```

```{r}
print(unpaired10)
```

```{r}
unpaired10.mean_diff <- mean_diff(unpaired10)

print(unpaired10.mean_diff)
```

```{r}
kable(unpaired10.mean_diff$boot_result |> select(-bootstraps))
```

```{r}
dabest_plot(unpaired10.mean_diff)
```

Supplementary Figure 5. Effects of low EPA concentrations on astrocyte viability.
b, b1. Cumming estimation plots showing astroglia viability (MTT assay) after 5 μM (b) or 10 μM (b1) EPA treatment. Left panel: Colored circles represent individual data points for control (Ctr) and EPA-treated groups from 6 independent experiments. Black circles and vertical lines indicate group means with 95% confidence intervals. Right panel: Floating plots show mean differences (EPA minus Ctr) for each experiment and the overall weighted mean difference (bottom). Circles represent the point estimate of the mean difference, with vertical lines indicating 95% confidence intervals. The shaded curve represents the resampled distribution of the effect size.

## 30 μM of EPA - Figure 5c

```{r}
unpaired30 <- load(df,
  x = Group, y = Measurement,
  idx = list(
    c("Ctr_1", "EPA_30μM_1"),
    c("Ctr_2", "EPA_30μM_2"),
    c("Ctr_3", "EPA_30μM_3"),
    c("Ctr_4", "EPA_30μM_4"),
    c("Ctr_5", "EPA_30μM_5"),
    c("Ctr_6", "EPA_30μM_6")
  ),
  minimeta = TRUE
)
```

```{r}
print(unpaired30)
```

```{r}
unpaired30.mean_diff <- mean_diff(unpaired30)

print(unpaired30.mean_diff)
```

```{r}
kable(unpaired30.mean_diff$boot_result |> select(-bootstraps))
```

```{r}
dabest_plot(unpaired30.mean_diff)
```

Figure 5. Effects of EPA on astrocyte viability.
c. Cumming estimation plot showing astroglia viability (MTT assay) after 30 μM EPA treatment. Left panel: Colored circles represent individual data points for control (Ctr) and EPA-treated groups from 6 independent experiments. Black circles and vertical lines indicate group means with 95% confidence intervals. Right panel: Floating plots show mean differences (EPA minus Ctr) for each experiment and the overall weighted mean difference (bottom). Circles represent the point estimate of the mean difference, with vertical lines indicating 95% confidence intervals. The shaded curve represents the resampled distribution of the effect size.

## 100 μM of Glutamate - Figure 3c

```{r}
unpaired_glut <- load(df_glut,
  x = Group, y = Measurement,
  idx = list(
    c("Ctr_1", "Glu_1"),
    c("Ctr_2", "Glu_2"),
    c("Ctr_3", "Glu_3"),
    c("Ctr_4", "Glu_4"),
    c("Ctr_5", "Glu_5"),
    c("Ctr_6", "Glu_6")
  ),
  minimeta = TRUE
)
```

```{r}
print(unpaired_glut)
```

```{r}
unpaired_glut.mean_diff <- mean_diff(unpaired_glut)

print(unpaired_glut.mean_diff)
```

```{r}
kable(unpaired_glut.mean_diff$boot_result |> select(-bootstraps))
```

```{r}
dabest_plot(unpaired_glut.mean_diff)
```

Figure 3. Effects of glutamate on astrocyte viability.
c. Cumming estimation plot showing astroglia viability (MTT assay) after 100 μM glutamate treatment. Left panel: Colored circles represent individual data points for control (Ctr) and glutamate-treated (Glu) groups from 6 independent experiments. Black circles and vertical lines indicate group means with 95% confidence intervals. Right panel: Floating plots show mean differences (Glu minus Ctr) for each experiment and the overall weighted mean difference (bottom). Circles represent the point estimate of the mean difference, with vertical lines indicating 95% confidence intervals. The shaded curve represents the resampled distribution of the effect size.




